## Clustering  

Libraries
```{r}
options(tidyverse.quiet=TRUE)
library(tidyverse)
library(cluster) #algorithms for clustering
library(factoextra) #visualization
library(dendextend) #viewing clustering dendograms
```


Read in data
```{r}
trucks = read_csv("trucks.csv")
str(trucks)
summary(trucks)
```
Preparing the data.
1. Remove missingness or impute missing values
2.  Scale the data

```{r}
trucks = trucks %>% select(-Driver_ID) 
str(trucks)
summary(trucks)
```

Basic data that we generate as an example (modification of code borrowed from https://www.tidymodels.org/learn/statistics/k-means/)

```{r}
set.seed(27)
centers = tibble(
  cluster = factor(1:2), #two clusters
  num_points = c(70, 120),  # number of points in each cluster
  Distance = c(5, 0),              # x1 coordinate of cluster centers
  Speeding = c(-1, 1)              # x2 coordinate of cluster centers
)

labelled_points = 
  centers %>%
  mutate(
    Distance = map2(num_points, Distance, rnorm), #randomly generate points for each cluster for x1
    Speeding = map2(num_points, Speeding, rnorm)  #randomly generate points for each cluster for x2
  ) %>% 
  select(-num_points) %>% 
  unnest(cols = c(Distance, Speeding))

ggplot(labelled_points, aes(Distance, Speeding, color = cluster)) +
  geom_point(alpha = 0.4) + theme_bw()
```


```{r}
kmeans_recipe = recipe(~ Distance + Speeding , trucks) 

trucks_dummy = kmeans_recipe %>% 
  step_dummy(all_nominal(), one_hot = TRUE) %>%
  step_scale(all_numeric()) %>%
  step_center(all_numeric()) 

trucks_dummy = prep(trucks_dummy, trucks) #prepares the recipe

trucks_cleaned = bake(trucks_dummy, trucks) #applies the recipe and yields a data frame
```

```{r}
#we have two data frames now, one scaled and dummied and one with just row-wise deletion of missingness
summary(trucks_cleaned)
summary(trucks)
```
Perform k-means clustering with a pre-specified number of clusters. We use the scaled and dummied data frame. 

```{r}
set.seed(412)
clusts = 
  tibble(k = 1:8) %>%
  mutate(
    kclust = map(k, ~kmeans(trucks_cleaned, .x)),
    tidied = map(kclust, tidy),
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, trucks_cleaned)
  )

clusts
```
Create relevant objects 
```{r}
clusters = 
  clusts %>%
  unnest(cols = c(tidied))

assignments = 
  clusts %>% 
  unnest(cols = c(augmented))

clusterings = 
  clusts %>%
  unnest(cols = c(glanced))
```

```{r}
p1 = 
  ggplot(assignments, aes(x = Distance, y = Speeding)) +
  geom_point(aes(color = .cluster), alpha = 0.8) + 
  facet_wrap(~ k)
p1
```
```{r}
ggplot(clusterings, aes(k, tot.withinss)) +
  geom_line() +
  geom_point()
```


```{r}
points = labelled_points %>% select(-cluster) #get rid of the cluster column

set.seed(123)
kclust = kmeans(points, centers = 2) #run k-means clustering with k = 3
kclust #view results
```


Let's try 4
```{r}
set.seed(412)
clusters = kmeans(trucks_cleaned, 4)
```

Attach cluster to dataset
```{r}
trucks = augment(clusters, trucks)
str(trucks)
```

Plots
```{r}
ggplot(trucks, aes(x=Distance,y=Speeding,color=factor(.cluster))) + geom_point() + facet_wrap(~factor(Speeding))
```




